#Johns Hopkins University on Coursera
##Practical Machine Learning Course Project

by Ivan Ribeiro

##Introduction

Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

##Analysis

The following assumes the raw data files were download and are stored in the Working Directory.

The following packages are required for this analysis and are here loaded.
```{r}
library(rpart)
library(rpart.plot)
library(e1071)
library(caret)
library(RColorBrewer)
library(randomForest)
library(rattle)
```

Reading raw files.
```{r}
train <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
test <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

To assure reproducibility:
```{r}
set.seed(88)
```

Here we will create a subset of the training data that will be used for cross validation.
```{r}
TrainPart <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
subtrain <- train[TrainPart, ]
subtest <- train[-TrainPart, ]
dim(subtrain)
dim(subtest)
```

At this point we are going to clean data. Variables with near zero variance, variables with mostly missing data an variables unuseful as predictors are removed.
```{r}
subtrain2 <- subtrain
for (i in 1:length(subtrain)) {
  if (sum(is.na(subtrain[ , i])) / nrow(subtrain) >= .75) {
    for (j in 1:length(subtrain2)) {
      if (length(grep(names(subtrain[i]), names(subtrain2)[j]))==1) {
        subtrain2 <- subtrain2[ , -j]
      }
    }
  }
}

subtrain3 <- subtrain2[,8:length(subtrain2)]

NZV <- nearZeroVar(subtrain3, saveMetrics = TRUE)
NZV
```

The same transformation is applied on testing sets
```{r}
clean1 <- colnames(subtrain3)
clean2 <- colnames(subtrain3[, -53]) #to remove classe
subtest <- subtrain3[clean1]
test <- test[clean2]
dim(subtest)
dim(test)
```

Now we will apply a Decision Tree (Fit1) and Random Forest (Fit2) and choose which model performs better. 
```{r}
Fit1 <- rpart(classe ~ ., data=subtrain3, method="class")
fancyRpartPlot(Fit1, main="Decision Tree", sub=" ")
```

```{r}
Fit2 <- randomForest(classe ~. , data=subtrain3)
Fit2
```

Now we will look at ou-ofsample errors applying the Models to subtest for Cross Validation
```{r}
Predict1 <- predict(Fit1, subtest, type = "class")
confusionMatrix(Predict1, subtest$classe)
Predict2 <- predict(Fit2, subtest, type = "class")
confusionMatrix(Predict2, subtest$classe)
```

As we decide that Random Forest is the most accurate model we will apply to the final test set
```{r}
Predict3 <- predict(Fit2, test)
Predict3
```
